### 2021-09-03 docs/paper/paper.tex
Add paper scaffold

\input{template/acm}
\input{template/bibtex}
\input{template/hyphenation}
\input{template/packages}
\input{template/commands}

\begin{document}

	\input{title}

	\input{sections/0_abstract}
	\keywords{mining software repositories, ecosystem call graphs, API usage, type analyzer, npm}

	\maketitle

	\input{sections/1_introduction}
	\input{sections/2_related_work}
	\input{sections/3_approach}
	\input{sections/4_data_collection}
	\input{sections/5_usage_mining}
	\input{sections/6_presentation}
	\input{sections/7_evaluation}
	\input{sections/8_conclusion}

	\input{acknowledgements}

	\input{template/bibliography}

	%\appendix

\end{document}
\endinput