### 2021-09-03 docs/paper/sections/1_introduction.tex

\section{Introduction}
\label{sec:introduction}

Foo.