### 2021-09-03 docs/paper/sections/2_related_work.tex

\section{Related work}
\label{sec:related_work}

Foo.