### 2021-09-03 docs/paper/sections/5_usage_mining.tex

\section{Usage mining}
\label{sec:usage_mining}

Foo!