### 2021-09-03 docs/paper/sections/7_evaluation.tex

\section{Evaluation}
\label{sec:evaluation}

Foo!