### 2021-09-03 docs/paper/sections/8_conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

Foo!