### 2021-09-03 docs/paper/sections/0_abstract.tex
Add abstract

\begin{abstract}

	In the dependency graph of a software ecosystem, \emph{downstream dependencies} are the nodes that depend on a package.
	Other than for upstream dependencies, solutions that provide individual package developers with this kind of information do not yet exist.
	This paper makes two major contributions:
	\begin{enumerate*}[label=(\roman*)]
		\item We propose an approach for mining usage data about npm packages from their downstream dependencies by utilizing a static type analyzer.
		\item We introduce a tool that allows package developers to survey usage samples of their APIs by incorporating the aggregated usage data into an integrated development environment.%
			\footnote{
				All artifacts are available on GitHub under the following URL: \url{https://github.com/LinqLover/downstream-repository-mining}
				\TODO{Insert name and use link to VSCE}
			}
	\end{enumerate*}
	We find that usage data from downstream dependency repositories are a promising source of information for mining software repositories and that they can be useful to support package developers in improving or extending their APIs.

\end{abstract}