### 2021-09-17 docs/paper/sections/4_dependency_collection.tex

\section{Downstream dependency collection}
\label{sec:dependency_collection}

In the first step, a list of repositories that depend on the target package has to be assembled.
As mentioned in \cref{sec:related_work/usage_samples}, many approaches start with a large downloaded corpus of unspecific repositories from the ecosystem which they then iterate over to filter the relevant repositories.
However, a major drawback of this approach is the high resource demand for both creating and traversing this corpus.
This approach is suited for analyzing repositories in large scale but is not conform to \cref{req1} due to the large footprint in terms of computational power and time.

As an alternative, we have decided to apply significant pre-filtering to the set of repositories before downloading it to the local machine.
For that, we have identified two types of data sources that are available in public clouds:

\begin{enumerate}[label=(\roman*)]
	\item Package repositories that maintain a doubly-connected edge list of all packages in an ecosystem depending on each other.
	\item Code search engines that index the source code of many repositories from OSS platforms.
		Using these search engines, we can query all repositories that declare a dependency on the target package in their package metadata file.
\end{enumerate}

Finally, we can download the source code for a small number of relevant repositories only.