### 2021-09-29 docs/paper/sections/5_usage_mining/patterns/JsxOpeningElement.tex

\begin{tikzpicture}
	\node [red] {JsxOpeningExpression}
		child [gray] {node [gray] {tagName}}
		child [gray] {node [gray, yshift = -0.5cm] {typeArguments}}
		child [gray] {node [gray] {attributes}};
\end{tikzpicture}
\caption[LoF entry]{
	Node pattern for a JSX opening element (as supported in React\protect\footnotemark{} or TypeScript\protect\footnotemark{}), such as in:
	\code{\lowlight{elem = }\uline{<Button \lowlight{color}="\lowlight{blue}"> \lowlight{Google</Button>}}\lowlight{;}}
}