### 2021-09-29 docs/paper/sections/8_conclusion.tex

\section{Conclusion and future work}
\label{sec:conclusion}

Foo!

future work:

* evaluation as future work:

	- rq2 - false negatives durch annotation ermitteln
	- rq3 - user study

* pattern analysis -> sample classification, potential for reuse
* dynamic analysis -> parameter/return value insights, coverage detection
* dependency classification/clustering through metrics (number of references, coupling)

minor:
* support more AST pattern (e.g. import())

STEHENGEBLIEBEN: Notizen hierfür, in slides übernehmen, dann weitere Notizen für Präsi anwenden. Heute nochmal durchsprehcne?