### 2021-09-29 docs/paper/sections/8_conclusion.tex
Remove notes committed by accident 😬

\section{Conclusion and future work}
\label{sec:conclusion}

Foo!