### 2021-10-08 docs/paper/paper.tex
Revise paper and apply feedback by @scheibel
- Restructure sections
- Shorten things
- Rephrasing
- Improve typesetting

\input{template/acm}
\input{template/bibtex}
\input{template/packages}
\input{template/patches}
\input{template/commands}
\input{template/lstlisting}
\input{hyphenation}
\input{crefnames}

% Colors
\colorlet{paperbackground}{white}
\input{colors}

\begin{document}

	\input{title}

	\input{sections/0_abstract}
	\keywords{mining software repositories, downstream dependencies, ecosystem call graphs, API usage, type analyzer}

	\maketitle
	\pagestyle{plain} % removes running headers from ACM template

	\input{sections/1_introduction}
	\input{sections/2_related_work}
	\input{sections/3_approach}
	\input{sections/4_implementation}
	\input{sections/5_evaluation}
	\input{sections/6_conclusion}

	\input{acknowledgements}

	\input{template/bibliography}

	%\appendix

\end{document}
\endinput