### 2021-10-08 docs/paper/sections/0_abstract.tex

\begin{abstract}

	In the dependency graph of a software ecosystem, \emph{downstream dependencies} are the nodes that depend on a package.
	Other than for upstream dependencies, solutions that provide individual package developers with this kind of information rarely exist to date.
	This paper makes two major contributions:
	\begin{enumerate*}[label=(\roman*)]
		\item We propose an approach for gathering downstream dependencies of a single package efficiently and extracting usage samples from them by utilizing a static type analyzer.
		\item We introduce a tool that allows npm package developers to survey usage samples of their APIs by incorporating the aggregated usage data into an IDE.%
			\footnote{
				All artifacts are available on GitHub under the following URL: \url{https://github.com/LinqLover/downstream-repository-mining}
				\TODO{Insert name and use link to VSCE}
			}
	\end{enumerate*}

	Our two methods yield over \num{8000} dependencies for popular packages at an average precision of \SI{98.7}{\percent} or \SI{68.4}{\percent}, resp., but tend to exclude unpopular dependencies.
	The extracted usage samples have a high precision but are unavailable for repositories with complex build configurations or metaprogramming patterns.
	Our tool is able to process up to 12 dependencies per minute while requiring about \SI{500}{\mega\byte} RAM in total and less than \SI{30}{\mega\byte} ROM per package.
	We find that usage data from downstream dependency repositories are a promising source of information for mining software repositories and that our approach could support package developers in improving or extending their APIs.

\end{abstract}